---
title: "books"
author: "Shaya Engelman"
date: "2023-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment promp:

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more 
than one author. For each book, include the title, authors, and two or three other attributes that you find 
interesting.
Take the information that you’ve selected about these three books, and separately create three files which 
store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”, 
“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you 
create each of these files “by hand” unless you’re already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into 
separate R data frames. Are the three data frames identical?
Your deliverable is the three source files and the R code. If you can, package your assignment solution up into 
an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files 
accessible from the web].



```{r libraries}
library(jsonlite)
library(tidyverse)
library(XML)
library(rvest)
```

```{r json}
books_json <- fromJSON("https://raw.githubusercontent.com/Shayaeng/Data607/main/books.json")
```

```{r xml}
books_xml_object <- xmlParse(getURL("https://raw.githubusercontent.com/Shayaeng/Data607/main/books.xml"))
books_xml <- xmlToDataFrame(books_xml_object)

```

```{r html}
books_html_raw <- read_html("https://raw.githubusercontent.com/Shayaeng/Data607/main/books.html")

table_data <- html_nodes(books_html_raw, "table") %>%
                html_table()
books_html <- as.data.frame(table_data)
```

```{r view tables}
(books_json)
(books_xml)
(books_html)
```

By viewing the tables we can see that there are differences between the different dataframes. To make them identical, I will unnest the json file, set the first row of the html file as the column names and make the author column look the same for all three. 

```{r}
#unnest list and combine json rows
books_json_unnested <- books_json %>%
  unnest(authors)
books_json_unnested <- books_json_unnested %>%
  rename(authors = author)

books_json_unnested <- books_json_unnested %>%
  group_by(title, series, goodreads, release_year, genre) %>%
  summarise(authors = paste(authors, collapse = ", ")) %>%
  ungroup()

#reorder the columns
books_json_unnested <- books_json_unnested[, c(1,2,6,3,4,5)]

#rename columns and get rid of extra row in html
colnames(books_html) <- books_html[1, ]
books_html <- books_html[-1, ]

#fix author cell in xml and html
books_xml[1, 3] <- gsub("Jordan", "Jordan, ", books_xml[1, 3])
books_html[1, 3] <- gsub("\\s{2,}", ", ", books_html[1, 3])

# Order books_xml
books_xml <- books_xml[order(books_xml$goodreads, decreasing = TRUE), ]

# Order books_html
books_html <- books_html[order(books_html$goodreads, decreasing = TRUE), ]

# Order books_json
books_json_unnested <- books_json_unnested[order(books_json_unnested$goodreads, decreasing = TRUE), ]

```

```{r}
reset_row_names <- function(df) {
  rownames(df) <- NULL  # Remove existing row names
  rownames(df) <- seq_len(nrow(df))  # Assign new row names as a numeric sequence
  return(df)
}

# Applying the function to each data frame
books_xml <- reset_row_names(books_xml)
books_html <- reset_row_names(books_html)
books_json_unnested <- reset_row_names(books_json_unnested)

# Convert json to df from tibble
books_json <- as.data.frame(books_json_unnested)
```

```{r}
# Check if all three data frames are identical
all_identical <- identical(books_xml, books_html) && identical(books_html, books_json)

# Display the result
print(paste("All three data frames are identical:", all_identical))

```

The three tables are now identical.
