---
title: "Week 9"
author: "Shaya Engelman"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment - Web APIs

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and 
transform it into an R DataFrame:

```{r }
library(tidyverse)
library(httr)
library(jsonlite)
library(dotenv)
```

```{r}
#load the api key in .env file
api_key <- Sys.getenv("api_key")
```

```{r}
#set period separately to make it easy to adjust
period <- 7

url <- sprintf("https://api.nytimes.com/svc/mostpopular/v2/viewed/%d.json", period)

response <- GET(url, query = list(`api-key` = api_key))

#check status
if (http_status(response)$category == "Success") {
  # Parse the JSON content
  data <- content(response, "text") %>%
    fromJSON(flatten = TRUE)
} else {
  #print error
  stop("Error: Unable to retrieve data from the API.")
}
```

```{r}
#convert to dataframe
df <- as.data.frame(data)
```

**Analyses**

I had some basic questions I wanted to collect from the list of most viewed articles I collected. I asked and plotted their answers.


```{r}
#plot the NYT section the top 20 articles appear in
ggplot(df, aes(x = results.section)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Frequency of Sections", x = "Section", y = "Frequency") +
  theme_minimal() +
  coord_flip()
```

```{r}
#plot the frequency of specific keywords
df <- df %>%
  mutate(keyword_appearance = grepl("Gaza|Israel", results.adx_keywords, ignore.case = TRUE))

#create a barplot of the frequency of rows with 'Gaza' or 'Israel' in the keywords
ggplot(df, aes(x = factor(keyword_appearance), fill = factor(keyword_appearance))) +
  geom_bar() +
  labs(title = "Frequency of 'Gaza' or 'Israel' in Keywords", x = "Israel/Gaza", y = "Frequency") +
  scale_x_discrete(labels = c("Does Not Appear", "Appears")) +
  scale_fill_manual(values = c("gray", "blue")) +  # Adjust the colors as needed
  theme_minimal() +
  guides(fill = FALSE)
```

